---
title: "Different Insect Visitors Shape Mango Harvests"
author: "Mia Fitzgerald and Angela McMahon"
date: "4/24/2026"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
library(knitr)
library(tidyverse)
library(readr)
library(janitor)

```


### Introduction 
Our data sets were generated from a field study which was conducted on six commercial mango orchards in Australia. The research was to assess the temporal and spacial variations in insect species visiting the orchard and their influence on the mango flowers. Specifically our first data set is looking at the fruit itself and the impact of insect visitation on reproductive success of the flower. This set has a sample size of 85. This set includes some different insect species along with an other category, the total number of insects that visited, what farm they are at, the distance, fruit size in different measurements and the harvest fruit set.  Our second data set examines the insect visitors during the night time with a sample size of 85 as well. This data set includes the farm and name as well, along with the time of day, weather conditions, and different insect species seen at those specific conditions. 
Link to both: https://datadryad.org/dataset/doi:10.5061/dryad.v6wwpzh7p 


### Data Import
After downloading the csv files from the link above. Imported the data to R studio, read them in, and stored under fruit_data and nocturnal_visits. The columns names had spaces so we cleaned them using the janitor package to make it compatible with R. Then we made the combined dataset using full_join because they were already in order with the same amount of observations. Join_by helped stop repeat column names and the total variables were renamed for clarity. 

```{r}
fruit_data <- read_csv("Fruit_set.csv") %>% clean_names() 
nocturnal_visits <- read_csv("Nocturnal_visits.csv") %>% clean_names() 

combined_insect_data <- fruit_data %>%
  full_join(nocturnal_visits, by = join_by(farm, distance)) %>%
  rename(nocturnal_total = total.y, fruit_total = total.x)

```


## Have an Exploratory Data Analysis section. a. Explore your variables. You don’t need to include every plot/table you try out, but include theclearest, most illustrating ones. b. Explore the primary/compound/foreign key variables in both data sets

```{r}
fruit_data %>%
  filter(distance %in% c(0, 50, 100, 150, 200)) %>%
  group_by(distance) %>%
  select(distance, total) %>%
  summarize(avg_total_insects = mean(total, na.rm = TRUE)) %>%
  kable()

```


```{r}
fruit_data %>%
  filter(!is.na(farm)) %>%
  ggplot(aes(x = total, y = weight_g)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Total Number of Insect Visitors on the Mangoes Weight", x = "total_insects")

```


```{r}
fruit_data %>%
  filter(!is.na(farm)) %>%
  ggplot(aes(x = total)) +
  geom_boxplot() +
  facet_wrap(~farm) +
  labs(title = "Total Amount of Insects at Each Farm", x = "total_insects")
  
```


```{r}
fruit_data%>%
  filter(!is.na(farm))%>%
  ggplot(aes(x=fruit_length_mm, y=fruit_breadth_mm, color=farm))+
  geom_point(alpha=0.7) + 
  facet_wrap(~farm) +
  labs(title = "Mango Size")
```


```{r}
fruit_data%>%
  summarize(avg_length=mean(fruit_length_mm, na.rm=TRUE),
            avg_width=mean(fruit_breadth_mm, na.rm=TRUE)) %>%
  kable()

```

# total insects in descending order -farm 6 saw no insects 
```{r}
nocturnal_visits%>%
  filter(!is.na(farm))%>%
  group_by(farm)%>%
  summarize(total_insect=sum(total))%>%
  arrange(desc(total_insect))%>%
  kable()
```


# intersting temp/humidity distribution over the farms - farm 6 shows very low temps and above average humidity 
```{r}
nocturnal_visits%>%
  filter(!is.na(farm))%>%
  rename(humidity=rh, temp.C=temp)%>%
  ggplot(aes(x=humidity, y=temp.C, color=farm))+
  geom_point(alpha=0.5)
```


# total insect counts seperated by distance from native veg- the closer the farm the more distibution 
```{r}
nocturnal_visits%>%
  filter(!is.na(farm))%>%
  ggplot(aes(x=total))+
  geom_histogram(binwidth=0.5)+
  facet_wrap(~distance)
```


# average number of insects visiting each marked distance-- the ones closest to the native veg are about 4x higher than any other distance
```{r}
nocturnal_visits%>%
  filter(!is.na(total))%>%
  group_by(distance)%>%
  summarize(avg_insect=round(mean(total),3))%>%
  kable()
```


# binned humidity levels based on the average seen humidity - very few cases have low humidity - most are about average
```{r}
nocturnal_visits%>%
  filter(!is.na(rh))%>%
  mutate(
    humidity= case_when(
      rh<65             ~ "low humidity",
      rh>=65 & rh<79     ~ "average humidity",
      rh>=79             ~ "high humidity"
    )
  )%>%ggplot(aes(x=humidity))+
  geom_bar()+
  labs(y="Number of Cases", x="Humidity Level", title="")



```


## 5 Research section 




## 6 Conclusion


